---
title: "Risk factors associated with COVID-19 mortality in patients with type 2 diabetes:
  a retrospective cohort study"
author: "Carlos Ballon-Salcedo & Kevin J Paez"
date: "`r Sys.Date()´"
format: 
  html:
    toc: true
    toc-depth: 4
    code-overflow: wrap
    code-fold: true
    code-tools: 
      source: "https://github.com/Carlos-Ballon/#####3"
    theme: 
     light: flatly
     dark: darkly
    highlight-style: github
editor: source
editor_options: 
  chunk_output_type: inline
bibliography: grateful-refs.bib
csl: vancouver.csl
---
# Set global knitr options

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Load packages

```{r}
# Packages
pacman::p_load(
  rio,
  here,
  reportfactory,
  rfextras,
  tidyverse,
  ggcorrplot,
  ggsci,
  ggpubr,
  ggbiplot,
  finalfit,
  gtsummary,
  flextable,
  ftExtra,
  broom,
  performance,
  lmtest,
  stats,
  moments,
  nortest,
  epiDisplay,
  car,
  Rtsne,
  factoextra,
  corrplot,
  grateful,
  patchwork,
  officer,
  mice
)
# My function scripts
#rfextras::load_scripts()
```
# Import data

```{r}
#clean_data <- import(here("data", "data.csv"))
```

# Set themes

## Define `gtsummary` theme

```{r}
my_gtsummary_theme <-
  list(
    "pkgwide-fn:pvalue_fun" = function(x)
      style_pvalue(x, digits = 2),
    "pkgwide-fn:prependpvalue_fun" = function(x)
      style_pvalue(x, digits = 2, prepend_p = TRUE),
    "tbl_summary-str:continuous_stat" = "{median} ({p25}, {p75})",
    "tbl_summary-str:categorical_stat" = "{n} ({p}%)",
    "tbl_summary-fn:percent_fun" = function(x)
      style_number(x, digits = 1, scale = 100),
    "tbl_summary-arg:missing" = "no"
  )

# Set a gtsummary theme
set_gtsummary_theme(my_gtsummary_theme, theme_gtsummary_compact())

# Set a gtsummary language
theme_gtsummary_language(language = "en")
```

## Define `flextable` theme

```{r}
my_flextable_theme <- function(x, bold_header = FALSE) {
  std_border <- fp_border(width = 1, color = "grey14")
  
  x <- border_remove(x)
  x <- hline_top(x, border = std_border, part = "header")
  x <- hline_bottom(x, border = std_border, part = "header")
  x <- bold(x, bold = bold_header, part = "header")
  x <- hline_bottom(x, border = std_border, part = "body")
  x <- align(x, align = "left", part = "body")
  x <- align(x, align = "center", part = "header")
  x <- font(x, part = "all", fontname = "Segoe UI")
  fix_border_issues(x, part = "all")
  autofit(x)
}
```

# Process data

## Remove missing values (\>10% by variable)

```{r}
# For imputation
vars_to_keep = c("glucose", "urea", "creatinine")

data_filtered = data |>
  dplyr::select(
    all_of(vars_to_keep),
    where(~ sum(is.na(.)) < 90)
  )
```

::: {style="text-align: justify"}
The following variables were removed: mcv, mch, hemoglobin, hematocrit, creatinine, urea, glucose, ph, anion_gap, sodio, potasio, cloro, calcio, fi_o2, pco2, hco3, lactate, antiparasitic.
:::

## Matching

The matchup will be with a 1:3 ratio match.

```{r}
# Matching for Causal Inference
data_match_object <- MatchIt::matchit(
  I(t2dm == "yes") ~ edad,
  data = data_filtered,
  exact = ~ sex,
  caliper = c(edad = 3),
  std.caliper = T,
  distance = "euclidean",
  ratio = 3
)

# Construct a matched dataset from a matchit object
data_match <- MatchIt::match.data(data_match_object, subclass = "matched_id")
```

## Recode and relevel (dictionary)

::: {style="text-align: justify"}
This subsection converts categorical variables into factors and recode/clean values using a data dictionary. The categorical variables will be saved as factor vectors, while some continuous/numeric variables will be categorized and converted into factors. Then, recode and change the factor levels using the `forcats` package. Also, the `finalfit` package will be utilized to label variables, while the `ftExtra` package will detect special characters and format the character columns as markdown text to be added to the flextable object. An alternate option is to utilize the `matchmaker` package for dictionary-based cleanup or `labelled` package to manipulate metadata. The disadvantage of categorizing continuous variables is that information is discarded. To address this, repeat the analysis on continuous variables to ensure there are no differences in the conclusion (sensitivity analysis).
:::

> 📝***NOTE:*** The markdown texts is by design a plain text

### Exposures and outcomes

::: {style="text-align: justify"}
Potential risk factors associated with mortality caused by COVID-19 in patients with type II diabetes mellitus (T2DM) are presented, each factor/variable is labeled with a legend that indicates its clinical importance, accuracy, and number of events. This is see in the **dictionary** script.
:::

-   Clinically important with enough evidence but with small number of events, shown as `\####`
-   Clinically important with enough evidence and enough number of events, shown as `\###`
-   Clinically important with limited evidence, shown as `\##`
-   Inconsistent or contradictory evidence and unconfirmed accuracy (self-reported) `\#`

```{r}
data_for_eda <- dictionary(data_match)
```

## Multiple Imputation with MICE
To adapt
https://stackoverflow.com/questions/53014141/mice-splitting-imputed-data-for-further-analysis

```{r}

colSums(is.na(input_data))

# Selecction of variables that may help to impute variables with missing data
input_data = 
  data_for_eda |>
    dplyr::select(
      # Demographic characteristics and clinical history
      edad,
      sex,
      t2dm,
      hta,
      # Signs and symptoms
      fever,
      tos,
      dolor_de_garganta,
      malestar_general,
      taquipnea,
      disnea, # Ok
      polidipysia,
      polifagia,
      # Vital signs
      frecuencia_respiratoria,
      frecuencia_cardiaca,
      p_a_sistolica,
      p_a_diastolica,
      # Laboratory findings
      wbc,
      neutrofilos,
      linfocitos,
      #nlr,
      platelets,
      #hemoglobina,
      #mcv,
      #mch,
      glucose,
      urea,
      creatinine,
      #lactate,
      # Blood gas findings
      saturacion_de_oxigeno,
      fio2,
      pafi,
      #pafi_cal,
      pao2,
      # Follow-up
      len_hosp_stay,
      # outcomes
      outcome
    )

## Imputation with mice. PMM = Predictive mean matching
data_imputated =
  mice(
    input_data,
    m = 20,
    method = c(
      "",
      "",
      "",
      "",
      "",
      "",
      "",
      "",
      "",
      "logreg",
      "",
      "",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      ""
    ),
    maxit = 20,
    seed = 500
    
  )

# For one variable
data_imputated$imp$len_hosp_stay

# # Complete
# data_match_eda_1 = complete(imput_b, 1)

# Adapt for design: separate for patients with diabetes and without diabetes

d.long <- mice::complete(data_imputated, "long", include = T)

d.long.t2dm <- d.long[which(d.long$t2dm == 'yes'),]
d.long.nont2dm <- d.long[which(d.long$t2dm == 'no'),]

data_imputated_t2dm <- as.mids(d.long.t2dm)
data_imputated_nont2dm <- as.mids(d.long.nont2dm)

data_imputated_t2dm$imp$creatinine
data_imputated_nont2dm$imp$creatinine
```

```{r fig.height=10, fig.width=10, fig.align='center'}
model_imputation_t2dm <- 
  with(imp.b_t2dm, glm(
    outcome ~ edad + hta + disnea + frecuencia_respiratoria + 
    frecuencia_cardiaca + p_a_sistolica + p_a_diastolica + 
    wbc + neutrofilos + linfocitos + platelets + glucose + 
    urea + creatinine + saturacion_de_oxigeno + fio2 + pafi + pao2,
    family = binomial(link = "logit")))

model_imputation_non <- 
  with(imp.b_nont2dm, glm(
    outcome ~ edad + hta + disnea + frecuencia_respiratoria + 
    frecuencia_cardiaca + p_a_sistolica + p_a_diastolica + 
    wbc + neutrofilos + linfocitos + platelets + glucose + 
    urea + creatinine + saturacion_de_oxigeno + fio2 + pafi + pao2,
    family = binomial(link = "logit")))

pool.imputation_non <- pool(model_imputation_non)
summary(pool.imputation_non)

# Visual check of model assumptions
performance::check_model(m3_non)

# Model performance
broom::glance(m3_non)

# Indices of model performance
performance::model_performance(m3_non)

# Check for multicollinearity
performance::check_collinearity(m3_non)
```




