---
title: "Risk factors associated with COVID-19 mortality in patients with type 2 diabetes:
  a retrospective cohort study"
author: "Carlos Ballon-Salcedo & Kevin J Paez"
date: "`r Sys.Date()´"
format: 
  html:
    toc: true
    toc-depth: 4
    code-overflow: wrap
    code-fold: true
    code-tools: 
      source: "https://github.com/Carlos-Ballon/#####3"
    theme: 
     light: flatly
     dark: darkly
    highlight-style: github
editor: source
editor_options: 
  chunk_output_type: inline
bibliography: grateful-refs.bib
csl: vancouver.csl
---
# Set global knitr options

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Load packages

```{r}
# Packages
pacman::p_load(
  rio,
  here,
  reportfactory,
  rfextras,
  tidyverse,
  ggcorrplot,
  ggsci,
  ggpubr,
  ggbiplot,
  finalfit,
  gtsummary,
  flextable,
  ftExtra,
  broom,
  performance,
  lmtest,
  stats,
  moments,
  nortest,
  epiDisplay,
  car,
  Rtsne,
  factoextra,
  corrplot,
  grateful,
  patchwork,
  officer,
  mice,
  knitr,
  survival,
  tibble,
  lubridate,
  ggsurvfit,
  gtsummary,
  tidycmprsk,
  ggmice
)
ezfun::set_ccf_palette("contrast")
# My function scripts
#rfextras::load_scripts()
```
# Import data

```{r}
#clean_data <- import(here("data", "data.csv"))
```

# Set themes

## Define `gtsummary` theme

```{r}
my_gtsummary_theme <-
  list(
    "pkgwide-fn:pvalue_fun" = function(x)
      style_pvalue(x, digits = 2),
    "pkgwide-fn:prependpvalue_fun" = function(x)
      style_pvalue(x, digits = 2, prepend_p = TRUE),
    "tbl_summary-str:continuous_stat" = "{median} ({p25}, {p75})",
    "tbl_summary-str:categorical_stat" = "{n} ({p}%)",
    "tbl_summary-fn:percent_fun" = function(x)
      style_number(x, digits = 1, scale = 100),
    "tbl_summary-arg:missing" = "ifany",
    "tbl_summary-arg:missing_text" = "Missing"
  )

# Set a gtsummary theme
set_gtsummary_theme(my_gtsummary_theme, theme_gtsummary_compact())

# Set a gtsummary language
theme_gtsummary_language(language = "en")
```

## Define `flextable` theme

```{r}
my_flextable_theme <- function(x, bold_header = FALSE) {
  std_border <- fp_border(width = 1, color = "grey14")
  
  x <- border_remove(x)
  x <- hline_top(x, border = std_border, part = "header")
  x <- hline_bottom(x, border = std_border, part = "header")
  x <- bold(x, bold = bold_header, part = "header")
  x <- hline_bottom(x, border = std_border, part = "body")
  x <- align(x, align = "left", part = "body")
  x <- align(x, align = "center", part = "header")
  x <- font(x, part = "all", fontname = "Segoe UI")
  fix_border_issues(x, part = "all")
  autofit(x)
}
```

# Process data

## Remove missing values (\>10% by variable)

```{r}
# For imputation
vars_to_keep = c("glucose", "urea", "creatinine")

data_filtered = data |>
  dplyr::select(
    all_of(vars_to_keep),
    where(~ sum(is.na(.)) < 90)
  )
```

::: {style="text-align: justify"}
The following variables were removed: mcv, mch, hemoglobin, hematocrit, creatinine, urea, glucose, ph, anion_gap, sodio, potasio, cloro, calcio, fi_o2, pco2, hco3, lactate, antiparasitic.
:::

## Matching

The matchup will be with a 1:3 ratio match.

```{r}
# Matching for Causal Inference
data_matched_object <- MatchIt::matchit(
  I(t2dm == "yes") ~ edad,
  data = data_filtered,
  exact = ~ sex,
  caliper = c(edad = 3),
  std.caliper = T,
  distance = "euclidean",
  ratio = 3
)

# Construct a matched dataset from a matchit object
data_matched_0 <- MatchIt::match.data(data_matched_object, subclass = "matched_id")
```

## Recode and relevel (dictionary)

::: {style="text-align: justify"}
This subsection converts categorical variables into factors and recode/clean values using a data dictionary. The categorical variables will be saved as factor vectors, while some continuous/numeric variables will be categorized and converted into factors. Then, recode and change the factor levels using the `forcats` package. Also, the `finalfit` package will be utilized to label variables, while the `ftExtra` package will detect special characters and format the character columns as markdown text to be added to the flextable object. An alternate option is to utilize the `matchmaker` package for dictionary-based cleanup or `labelled` package to manipulate metadata. The disadvantage of categorizing continuous variables is that information is discarded. To address this, repeat the analysis on continuous variables to ensure there are no differences in the conclusion (sensitivity analysis).
:::

> 📝***NOTE:*** The markdown texts is by design a plain text

### Exposures and outcomes

::: {style="text-align: justify"}
Potential risk factors associated with mortality caused by COVID-19 in patients with type II diabetes mellitus (T2DM) are presented, each factor/variable is labeled with a legend that indicates its clinical importance, accuracy, and number of events. This is see in the **dictionary** script.
:::

-   Clinically important with enough evidence but with small number of events, shown as `\####`
-   Clinically important with enough evidence and enough number of events, shown as `\###`
-   Clinically important with limited evidence, shown as `\##`
-   Inconsistent or contradictory evidence and unconfirmed accuracy (self-reported) `\#`

```{r}
data_for_eda <- dictionary(data_matched_0)
```
## Exploratory Data Analysis (EDA). T2DM is the exposure variable.

### Categorical variables

```{r echo=TRUE, include=FALSE}
# Factor selection
categorical <- data_for_eda |>
  dplyr::select(where(is.factor), -matched_id)

# Multiple tables (relative frequencies) with lapply
lapply(categorical, function(x)
  round(prop.table(table(x)) * 100, 1))

# Relative frequencies with tidyr and dplyr
categorical_pivot <- categorical |>
  tidyr::pivot_longer(everything(), names_to = "Variable", values_to = "Valor") |>
  dplyr::group_by(Variable, Valor) |>
  dplyr::summarise(n = n()) |>
  tidyr::drop_na() |>
  dplyr::mutate(Porcentaje = round(n / sum(n) * 100, 1))
```

```{r}
# Variables with relative frequency less than 2.5%
excluded_var <- categorical_pivot |>
  dplyr::filter(Porcentaje < 2.5) |>
  dplyr::distinct(Variable) |>
  dplyr::pull()
```

```{r eval=FALSE}
# Multiple tables by T2DM
lapply(categorical, function(x)
  table(x, categorical$t2dm))
```

### Numerical variables

::: {style="text-align: justify"}
The distribution of continuous/numeric variables can be visualized using histograms, frequency polygons, Q-Q plots, box plots, violin plots, jitter plots, and Sina plots. A variable is considered continuous if it can assume any value from an infinite set of ordered values within an interval. A histogram shows the distribution of a continuous variable and differs from a bar chart by grouping data into continuous intervals. It also examines the distribution of a continuous variable broken down by a categorical variable (categorical outcome). A density plot displays the density instead of the count, which is the standardized count so that the area under each frequency polygon is 1 or 100%. In addition to the histogram and the frequency polygon, the probability distribution of a continuous variable can be obtained from the **density function**. This function is defined in terms of probability and would construct a frequency polygon if we had an infinite set of data, resulting in a continuous curve.
:::

```{r results='hide'}
# Selection of numerical variables
numerical <- data_match_eda |>
  dplyr::select(where(is.numeric))

# Summary statistics
lapply(numerical, function(x)
  summary(x))
```

```{r fig.height=10, fig.width=10, fig.align='center'}
# Variable groups
dem_numerical <- numerical |>
  dplyr::select(edad:p_a_diastolica)

hemo_numerical <- numerical |>
  dplyr::select(wbc:platelets)

gas_numerical <- numerical |>
  dplyr::select(saturacion_de_oxigeno:pao2)

# Formal and visual exploration
total_plots(dem_numerical)
```

### Numerical variables in survivors and non-survivor

::: {style="text-align: justify"}
In this part, density plots will be used to illustrate the data grouped by the outcome. Other plots will not be included in the analysis, as a table with more detailed information will be presented. The questions we will ask ourselves are: Does the data appear to be normally distributed? and are the variances between groups equal?
:::

```{r}
# Selection of numerical variables of patients with T2DM
surv_numerical_t2dm <- data_match_eda |>
  dplyr::select(where(is.numeric), outcome, t2dm) |>
  dplyr::filter(t2dm == "yes")

# # Selection of numerical variables of patients without T2DM
non_numerical_not2dm <- data_match_eda |>
  dplyr::select(where(is.numeric), outcome, t2dm) |>
  dplyr::filter(t2dm == "no")
```

```{r eval=FALSE}
# Summary statistics of survivors
lapply(surv_numerical_t2dm, function(x)
  summary(x))

# Summary statistics of non-survivors
lapply(non_numerical_not2dm, function(x)
  summary(x))
```

```{r fig.height=15, fig.width=18, fig.align = 'center'}
# Exploration of age by group
p1 <- group_stat_table_plot(surv_numerical_t2dm, "frecuencia_cardiaca", "outcome")

# Exploration of white-cells by group
p2 <- group_stat_table_plot(surv_numerical_t2dm, "pafi", "outcome")

# Exploration of neutrophils by group
p3 <- group_stat_table_plot(surv_numerical_t2dm, "pafi_cal", "outcome")

# Exploration of lymphocytes by group
p4 <- group_stat_table_plot(surv_numerical_t2dm, "pao2", "outcome")

p1 + p2 + p3 + p4 + plot_annotation(tag_levels = 'A')
```

### Correlation matrix (multicollinearity)

::: {style="text-align: justify"}
The correlation between independent variables will be explored in order to identify multicollinearity. The numeric variables are different, with the exception of white-cells and the PaO~2~:FiO~2~ ratio. In the case of using white-cells and neutrophils or lymphocytes, it's possible that the information provided by white-cells may be redundant and generate biased estimates. The same applies to the PaO~2~:FiO~2~ ratio and the FiO~2~. Given the clinical relevance of the PaO₂:FiO₂ ratio, it is preferable to work with this ratio rather than the PaO₂ alone.
:::

> 📝***NOTE:*** ggplot objects can't process markdown text

```{r fig.height=9, fig.width=9, fig.align='center'}
# Selection of numerical variables
data_num = data_match_eda |>
  dplyr::select(where(is.numeric), -weights) |>
  na.omit()

# Rename variables
cor_data = data_num |> dplyr::rename("Age" = edad, 
"Respiratory rate" = frecuencia_respiratoria,
  "Heart rate" = frecuencia_cardiaca,
  "SBP" = p_a_sistolica,
  "DBP" = p_a_diastolica,
  "White-cells" = wbc,
  "Neutrophils" = neutrofilos,
  "Lymphocytes" = linfocitos,
  "NLR" = nlr,
  "Platelets" = platelets,
  "SaO2" = saturacion_de_oxigeno,
  "FiO2" = fio2,
  "PaO2:Fio2 ratio" = pafi,
  "Calculated\nPaO2:Fio2 ratio" = pafi_cal,
  "PaO2" = pao2,
  "Follow-up time (days)" = len_hosp_stay
)

# Correlation matrix
corr <- round(cor(cor_data), 1)

# Color visualization
FS1 <- my_ggcorrplor(corr)

# View
FS1

# Gray visualization
FS1_grey <- my_ggcorrplor_grey(corr)
```
# Produce outputs

## Variable selection to analyze

::: {style="text-align: justify"}
The variables with not enough events (\<2.5%) were excluded from the analysis. These were alcoholism, antibiotics, asma_bronquial, cancer, disf_multiorg, dislip, ecv, epc, erc, f_hepatica_aguda, f_multiorganica, hemodialisis, hiv, immunsupressive_dis, perdida_de_peso, polidipysia, polifagia, poliuria, sensorio, and smoker. The variables ac_renal_failure, sepsis and shock_septico were also excluded due to possible misclassification bias.
:::

```{r}
# Select included variables
data_matched <- var_selec_analysis(data_for_eda)
```

```{r}
# Check included variables
included_var <- colnames(data_match)

# Check excluded variables
dplyr::setdiff(excluded_var, included_var)
#unique(excluded_var, included_var)
```

## Table 1. Demographic and clinical characteristics of patients

```{r}
# Demographic characteristics and clinical history
tbl_1.1 <- data_matched |>
  tbl_summary(
    include = c(edad:hta),
    by = t2dm,
    percent = "column",
    statistic = list(edad ~ "{mean} ({sd})"),
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_overall() |>
  add_p(edad ~ "t.test",
        test.args = all_tests("t.test") ~ list(var.equal = TRUE)) |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})",
                stat_0 = "**All patients** (n = {N})",
                p.value = "**p value**") |>
  modify_spanning_header(c(stat_1, stat_2) ~ "**Type 2 Diabetes**")

# Signs and symptoms
tbl_1.2 <- data_matched |>
  tbl_summary(
    include = c(fever:dolor_abdominal, t2dm),
    by = t2dm,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_overall() |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")

# Vital signs
tbl_1.3 <- data_matched |>
  tbl_summary(
    include = c(frecuencia_respiratoria:p_a_diastolica, t2dm),
    by = t2dm,
    percent = "column",
    statistic = list(frecuencia_cardiaca ~ "{mean} ({sd})"),
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_overall() |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")
```

```{r}
# Stack tables
table_1 = tbl_stack(
  list(tbl_1.1, tbl_1.2, tbl_1.3),
  group_header = c(
    "Demographic characteristics and clinical history",
    "Signs and symtoms",
    "Vital signs"
  ),
  quiet = TRUE
) |>
  modify_caption(
    "**Table 1**. Demographic and clinical characteristics of the patients"
    )

# Convert gtsummary object to a flextable object and format character columns
flex_table_1 <- table_1 |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
table_1

```

## Table 2. Laboratory findings and treatment of patients

```{r}
# Laboratory findings
tbl_2.1 <- data_matched |>
  tbl_summary(
    include = c(wbc:creatinine, t2dm),
    by = t2dm,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_overall() |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})",
                stat_0 = "**All patients** (n = {N})",
                p.value = "**p value**") |>
  modify_spanning_header(all_stat_cols(stat_0 = FALSE) ~ "**Type 2 Diabetes**")

# Blood gas findings
tbl_2.2 <- data_matched |>
  tbl_summary(
    include = c(saturacion_de_oxigeno:pao2.c, t2dm),
    by = t2dm,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_overall() |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")

# Treatments
tbl_2.3 <- data_matched |>
  tbl_summary(
    include = c(corticoides:pronacion, 
                #outcome, 
                t2dm),
    by = t2dm,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_overall() |>
  add_p() |>
  bold_p(t = 0.05) |> 
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")
```

```{r}
# Stack tables
table_2 = tbl_stack(
  list(tbl_2.1, tbl_2.2, tbl_2.3),
  group_header = c("Laboratory findings", "Blood gas findings", "Treatment"),
  quiet = TRUE
) |>
  modify_caption("Table 2. Laboratory findings and treatment of patients")

# Convert gtsummary object to a flextable object and format character columns
flex_table_2 <- table_2 |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
table_2
```

## Table 3. Demographic and clinical characteristics of patients by T2DM

### Patients with T2DM

```{r}
# Demographic characteristics and clinical history
tbl_3.1 <-
  data_matched |> 
  dplyr::filter(t2dm == "yes") |>
  tbl_summary(
    include = c(edad:obesity, hta, outcome),
    by = outcome,
    percent = "column",
    statistic = list(edad ~ "{mean} ({sd})"),
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})",
                p.value = "**p value**")

# Signs and symptoms
tbl_3.2 <-
  data_matched |> 
  dplyr::filter(t2dm == "yes") |>
  tbl_summary(
    include = c(fever:dolor_abdominal, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")

# Vital signs
tbl_3.3 <-
  data_matched |> 
  dplyr::filter(t2dm == "yes") |>
  tbl_summary(
    include = c(frecuencia_respiratoria:p_a_diastolica.c, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")
```
### Patients without T2DM

```{r}
# Demographic characteristics and clinical history
tbl_3.4 <-
  data_matched |>
  dplyr::filter(t2dm == "no") |>
  tbl_summary(
    include = c(edad:obesity, hta, outcome),
    by = outcome,
    percent = "column",
    statistic = list(edad ~ "{mean} ({sd})"),
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})",
                p.value = "**p value**")

# Signs and symptoms
tbl_3.5 <-
  data_matched |> 
  dplyr::filter(t2dm == "no") |>
  tbl_summary(
    include = c(fever:dolor_abdominal, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")

# Vital signs
tbl_3.6 <-
  data_matched |> 
  dplyr::filter(t2dm == "no") |>
  tbl_summary(
    include = c(frecuencia_respiratoria:p_a_diastolica.c, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")
```

```{r}
# Merge tables
tbl_3_p1 <- tbl_merge(
  tbls = list(tbl_3.1, tbl_3.4),
  tab_spanner = c("**Patients with T2DM**", "**Patients without T2DM**")
)

tbl_3_p2 <- tbl_merge(tbls = list(tbl_3.2, tbl_3.5)) |>
  modify_spanning_header(everything() ~ NA_character_)

tbl_3_p3 <- tbl_merge(tbls = list(tbl_3.3, tbl_3.6)) |>
  modify_spanning_header(everything() ~ NA_character_)

# Stack tables
table_3 = tbl_stack(
  list(tbl_3_p1, tbl_3_p2, tbl_3_p3),
  group_header = c(
    "Demographic characteristics and clinical history",
    "Signs and symtoms",
    "Vital signs"
  ),
  quiet = TRUE
) |> 
  modify_caption(
    "**Table 3**. Demographic and clinical characteristics of the patients by T2DM")

# Convert gtsummary object to a flextable object and format character columns
flex_table_3 <- table_3 |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
table_3
```
## Table 4. Laboratory findings and treatment of patients by T2DM

### Patients with T2DM

```{r}
# Laboratory findings
tbl_4.1 <-
  data_matched |> 
  dplyr::filter(t2dm == "yes") |>
  tbl_summary(
    include = c(wbc:creatinine, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})",
                p.value = "**p value**")

# Blood gas findings
tbl_4.2 <-
  data_matched |> 
  dplyr::filter(t2dm == "yes") |>
  tbl_summary(
    include = c(saturacion_de_oxigeno:pao2.c, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")

# Treatment
tbl_4.3 <-
  data_matched |> 
  dplyr::filter(t2dm == "yes") |>
  tbl_summary(
    include = c(corticoides:pronacion, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")
```

### Patients without T2DM

```{r}
# Laboratory findings
tbl_4.4 <-
  data_matched |> 
  dplyr::filter(t2dm == "no") |>
  tbl_summary(
    include = c(wbc:creatinine, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})",
                p.value = "**p value**")

# Blood gas findings
tbl_4.5 <-
  data_matched |> 
  dplyr::filter(t2dm == "no") |>
  tbl_summary(
    include = c(saturacion_de_oxigeno:pao2.c, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")

# Treatment
tbl_4.6 <-
  data_matched |> 
  dplyr::filter(t2dm == "no") |>
  tbl_summary(
    include = c(corticoides:pronacion, outcome),
    by = outcome,
    percent = "column",
    digits = list(all_continuous() ~ c(1, 1))
  ) |>
  add_p() |>
  bold_p(t = 0.05) |>
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})")
```

```{r}
# Merge tables
tbl_4_p1 <- tbl_merge(
  tbls = list(tbl_4.1, tbl_4.4),
  tab_spanner = c("**Patients with T2DM**", "**Patients without T2DM**")
)

tbl_4_p2 <- tbl_merge(tbls = list(tbl_4.2, tbl_4.5)) |>
  modify_spanning_header(everything() ~ NA_character_)

tbl_4_p3 <- tbl_merge(tbls = list(tbl_4.3, tbl_4.6)) |>
  modify_spanning_header(everything() ~ NA_character_)

# Stack tables
table_4 = tbl_stack(
  list(tbl_4_p1, tbl_4_p2, tbl_4_p3),
  group_header = c("Laboratory findings", "Blood gas findings", "Treatment"),
  quiet = TRUE
) |>
  modify_caption("Table 4. Laboratory findings and treatment of patients")


# Convert gtsummary object to a flextable object and format character columns
flex_table_4 <- table_4 |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
table_4
```

## Table 5. Cox Regression

::: {style="text-align: justify"}
In the univariate analysis, variables that were self-reported or did not have enough events were eliminated; this will help to avoid *overfitting* in the subsequent multivariable analysis. Other variables such as sex, smoker, alcoholism, dislip, obesity, malestar_general, anosmia, disgeusia, emesis, diarrea, poliuria, polidipysia, sensorio, p_a_sistolica, p_a_diastolica, platelets.c, pafi_cal.c, antibiotics, corticoides, anticoagulantes, pronacion, showed *no differences* between groups in the bivariate analysis (Tables 4 and 5).
:::

| Self-reported                                                                                                                                                              | Not enough event (\<10)                                                                                                                                                                                                                               |
|------------------------------------|------------------------------------|
| dolor_de_garganta, malestar_general, headache, anosmia, disgeusia, diarrea, emesis, astenia, dolor_abdominal, perdida_de_peso, sensorio, poliuria, polidipysia, polifagia. | smoker, alcoholism, dislip, ecv, cancer, hiv, immunsupressive_dis, erc, hemodialisis, asma_bronquial, epc, shock_septico, disf_multiorg, f_hepatica_aguda, f_multiorganica, perdida_de_peso, poliuria, polidipysia, polifagia, sensorio, antibiotics. |

: Not analyzed variables

```{r}
# Patients with T2DM
data_uv_dm <- var_selec_uv(data_matched, "yes")

# Patients without T2DM
data_uv_nondm <- var_selec_uv(data_matched, "no")
```
```{r}
# All patients
data_uv = 
  data_matched |> 
  dplyr::select(
      # Demographic characteristics and clinical history
      edad.c,
      sex,
      hta,
      obesity,
      
      # Signs and symptoms
      fever,
      tos,
      taquipnea,
      disnea,
      
      # Vital signs
      frecuencia_cardiaca.c,
      frecuencia_respiratoria.c,
      p_a_sistolica.c,
      p_a_diastolica.c,
      
      # Laboratory findings
      wbc.c,
      neutrofilos.c,
      linfocitos.c,
      nlr.c,
      platelets.c,
      creatinine,
      urea,
      glucose.c,
      saturacion_de_oxigeno.c,
      fio2.c,
      pafi.c,
      pao2.c,
      
      # Treatment
      corticoides,
      anticoagulantes,
      antipaludicos,
      pronacion,
      
      # outcomes
      outcome,
      outcome_ph,
      len_hosp_stay
    ) |>
  na.omit()
```

### Univariate models - No multiple data imputation

> 📝***NOTE:*** The variables included in the univariate regression analysis were selected based on the results of the bivariate analysis.

#### Patients with T2DM

```{r}
table_s1.1_uv <- data_uv_dm |>
  tbl_uvregression(
    include = c(edad.c:pronacion),
    y = Surv(len_hosp_stay, outcome_ph),
    method = coxph,
    exponentiate = TRUE,
    conf.int = TRUE,
    hide_n = TRUE,
    add_estimate_to_reference_rows = FALSE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      sex ~ "Sex",
      hta ~ "Hypertension",
      obesity ~ "Obesity",
      fever ~ "Fever",
      tos ~ "Dry cought",
      taquipnea ~ "Tachypnea",
      disnea ~ "Dyspnea",
      frecuencia_cardiaca.c ~ "Heart rate (BPM)",
      frecuencia_respiratoria.c ~ "Respiratory rate (BPM)",
      p_a_sistolica.c ~ "SBP (mmHg)",
      p_a_diastolica.c ~ "DBP (mmHg)",
      wbc.c ~ "White-cells (×10^−9^/L)",
      neutrofilos.c ~ "Neutrophils (×10^−9^/L)",
      linfocitos.c ~ "Lymphocytes (×10^−9^/L)",
      nlr.c ~ "NLR",
      platelets.c ~ "Platelets (×10^−9^/L)",
      creatinine ~ "Creatinine (mg/dL)",
      urea ~ "Urea (mg/dL)",
      glucose.c ~ "Glucose (mg/dL)",
      saturacion_de_oxigeno.c ~ "SaO~2~",
      fio2.c ~ "FiO~2~ (%)",
      pafi.c ~ "PaO~2~:FiO~2~ ratio",
      pao2.c ~ "PaO~2~",
      corticoides ~ "Corticosteroids",
      anticoagulantes ~ "Anticoagulants",
      antipaludicos ~ "Antimalarials",
      pronacion ~ "Pronation"
    )
  ) |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Univariate HR**", p.value = "**p value**")

```

#### Patients without T2DM

```{r}
table_s1.2_uv <- data_uv_nondm |>
  tbl_uvregression(
    include = c(edad.c:pronacion),
    y = outcome,
    method = glm,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    conf.int = TRUE,
    hide_n = TRUE,
    add_estimate_to_reference_rows = FALSE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      sex ~ "Sex",
      hta ~ "Hypertension",
      obesity ~ "Obesity",
      fever ~ "Fever",
      tos ~ "Dry cought",
      taquipnea ~ "Tachypnea",
      disnea ~ "Dyspnea",
      frecuencia_cardiaca.c ~ "Heart rate (BPM)",
      frecuencia_respiratoria.c ~ "Respiratory rate (BPM)",
      p_a_sistolica.c ~ "SBP (mmHg)",
      p_a_diastolica.c ~ "DBP (mmHg)",
      wbc.c ~ "White-cells (×10^−9^/L)",
      neutrofilos.c ~ "Neutrophils (×10^−9^/L)",
      linfocitos.c ~ "Lymphocytes (×10^−9^/L)",
      nlr.c ~ "NLR",
      platelets.c ~ "Platelets (×10^−9^/L)",
      creatinine ~ "Creatinine (mg/dL)",
      urea ~ "Urea (mg/dL)",
      glucose.c ~ "Glucose (mg/dL)",
      saturacion_de_oxigeno.c ~ "SaO~2~",
      fio2.c ~ "FiO~2~ (%)",
      pafi.c ~ "PaO~2~:FiO~2~ ratio",
      pao2.c ~ "PaO~2~",
      corticoides ~ "Corticosteroids",
      anticoagulantes ~ "Anticoagulants",
      antipaludicos ~ "Antimalarials",
      pronacion ~ "Pronation"
    )
  ) |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Univariate HR**", p.value = "**p value**")
```
#### All Patients

```{r}
table_s1.3_uv <- data_uv |>
  tbl_uvregression(
    include = c(edad.c:pronacion),
    y = outcome,
    method = glm,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    conf.int = TRUE,
    hide_n = TRUE,
    add_estimate_to_reference_rows = FALSE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      sex ~ "Sex",
      hta ~ "Hypertension",
      obesity ~ "Obesity",
      fever ~ "Fever",
      tos ~ "Dry cought",
      taquipnea ~ "Tachypnea",
      disnea ~ "Dyspnea",
      frecuencia_cardiaca.c ~ "Heart rate (BPM)",
      frecuencia_respiratoria.c ~ "Respiratory rate (BPM)",
      p_a_sistolica.c ~ "SBP (mmHg)",
      p_a_diastolica.c ~ "DBP (mmHg)",
      wbc.c ~ "White-cells (×10^−9^/L)",
      neutrofilos.c ~ "Neutrophils (×10^−9^/L)",
      linfocitos.c ~ "Lymphocytes (×10^−9^/L)",
      nlr.c ~ "NLR",
      platelets.c ~ "Platelets (×10^−9^/L)",
      creatinine ~ "Creatinine (mg/dL)",
      urea ~ "Urea (mg/dL)",
      glucose.c ~ "Glucose (mg/dL)",
      saturacion_de_oxigeno.c ~ "SaO~2~",
      fio2.c ~ "FiO~2~ (%)",
      pafi.c ~ "PaO~2~:FiO~2~ ratio",
      pao2.c ~ "PaO~2~",
      corticoides ~ "Corticosteroids",
      anticoagulantes ~ "Anticoagulants",
      antipaludicos ~ "Antimalarials",
      pronacion ~ "Pronation"
    )
  ) |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Univariate HR**", p.value = "**p value**")
```

```{r}
# Stack tables
table_uv = tbl_merge(
  list(table_s1.1_uv, table_s1.2_uv),
  tab_spanner = c("**Patients with T2DM**", "**Patients without T2DM**"
  )
) |>
  modify_caption("Table S1. Univariate Cox proportional-hazard regression by T2DM")

# Convert gtsummary object to a flextable object and format character columns
flex_table_s1 <- table_uv |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
flex_table_s1
```

## Figures from survival analysis

### Survival curves

```{r}
s1 <- survfit(Surv(len_hosp_stay, outcome_ph) ~ 1, data = data_matched)
str(s1)
```

```{r}
survfit2(Surv(len_hosp_stay, outcome_ph) ~ 1, data = data_uv_nondm) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +
  add_confidence_interval() +
  add_risktable()
```

```{r}
summary(survfit(Surv(len_hosp_stay, outcome_ph) ~ 1, data = data_matched), times = 10)
```
### Median survival time

```{r}
survfit(Surv(len_hosp_stay, outcome_ph) ~ 1, data = data_matched) %>% 
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median survival (95% CI)**"
  )

survdiff(Surv(len_hosp_stay, outcome_ph) ~ t2dm, data = data_matched)
```
```{r}
coxph(Surv(len_hosp_stay, outcome_ph) ~ t2dm, data = data_matched) %>% 
  tbl_regression(exp = TRUE) 
```

### Comparing patients with diabetes versus patients without diabetes

```{r}
# Raw differences
 
survdiff(Surv(len_hosp_stay, outcome_ph) ~ t2dm, data = data_matched)

# Quantify effect
coxph(Surv(len_hosp_stay, outcome_ph) ~ t2dm, data = data_matched)

```

## Comparison of survival curves

```{r}
os_age <-  survfit(Surv(len_hosp_stay, outcome_ph) ~ edad.c, data = data_uv_dm)
ggsurvplot( 
  os_age,
  data = data_uv_dm,
  size = 1, 
  linetype = "strata",
  #conf.int = T,
  surv.scale = "percent",  
  break.time.by = 10, 
  xlab = "Follow-up days",
  ylab= "Survival Probability",
  pval = T,
  pval.coord = c(40,.91),
  #risk.table = T,
  legend.title = "Age (years)",
  legend.labs = c("Less than 61", "Equal or greater than 61"),
  font.legend = 14,
  palette = c("#E7B800","#3E606F")
  #surv.median.line = "hv", 
  #ggtheme = theme_light()
)

os_age_dm = survfit(Surv(len_hosp_stay, outcome_ph) ~ edad.c, data = data_uv_dm)

os_dm2 = survfit(Surv(len_hosp_stay, outcome_ph) ~ t2dm, data = data_matched)

os_neut_dm = survfit(Surv(len_hosp_stay, outcome_ph) ~ neutrofilos.c, data = data_uv_dm)
os_nlr_dm = survfit(Surv(len_hosp_stay, outcome_ph) ~ nlr.c, data = data_uv_dm)
os_pao_dm = survfit(Surv(len_hosp_stay, outcome_ph) ~ pao2.c, data = data_uv_dm)
os_sao_dm = survfit(Surv(len_hosp_stay, outcome_ph) ~ saturacion_de_oxigeno.c, data = data_uv_dm)

plot_os = 
  ggsurvplot(os_dm2,
  data = data_matched,
  size = 1, 
  linetype = "strata",
  break.time.by = 10, 
  xlab = "Follow-up days",
  ylab = "Survival Probability",
  pval = T,
  legend.title = "Age (years)",
  #legend.labs = c("Less than 61", "61 or greater"),
  palette = c("#0073C2FF", "#EFC000FF"),
  font.legend = 13
  )
```
### Multivariable models

#### Complete multivariable for patients with T2DM

```{r}
# T2DM model
full_multivariable_dm <- glm(
  outcome ~
    edad.c + sex + hta + obesity + fever + tos + taquipnea + disnea +
    frecuencia_cardiaca.c + frecuencia_respiratoria.c + p_a_sistolica.c +
    p_a_diastolica.c + wbc.c + neutrofilos.c + linfocitos.c + nlr.c +
    platelets.c + saturacion_de_oxigeno.c + fio2.c + pafi.c + pao2.c +
    corticoides + anticoagulantes + antipaludicos + pronacion,
  data = data_uv_dm,
  family = binomial(link = "logit")
) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      sex ~ "Sex",
      hta ~ "Hypertension",
      obesity ~ "Obesity",
      fever ~ "Fever",
      tos ~ "Dry cought",
      taquipnea ~ "Tachypnea",
      disnea ~ "Dyspnea",
      frecuencia_cardiaca.c ~ "Heart rate (BPM)",
      frecuencia_respiratoria.c ~ "Respiratory rate (BPM)",
      p_a_sistolica.c ~ "SBP (mmHg)",
      p_a_diastolica.c ~ "DBP (mmHg)",
      wbc.c ~ "White-cells (×10^−9^/L)",
      neutrofilos.c ~ "Neutrophils (×10^−9^/L)",
      linfocitos.c ~ "Lymphocytes (×10^−9^/L)",
      nlr.c ~ "NLR",
      platelets.c ~ "Platelets (×10^−9^/L)",
      saturacion_de_oxigeno.c ~ "SaO~2~",
      fio2.c ~ "FiO~2~ (%)",
      pafi.c ~ "PaO~2~:FiO~2~ ratio",
      pao2.c ~ "PaO~2~",
      corticoides ~ "Corticosteroids",
      anticoagulantes ~ "Anticoagulants",
      antipaludicos ~ "Antimalarials",
      pronacion ~ "Pronation"
    )
  ) |>
  bold_p(t = 0.05) |>
  add_vif() |>
  modify_header(estimate = "**Multivariable OR**", 
                p.value = "**p value**", 
                aGVIF ~ "**aGVIF**")
```

```{r fig.height=10, fig.width=10, fig.align='center'}
# Model
m1_dm = glm(
  outcome ~
    edad.c + sex + hta + obesity + fever + tos + taquipnea + disnea +
    frecuencia_cardiaca.c + frecuencia_respiratoria.c + p_a_sistolica.c +
    p_a_diastolica.c + wbc.c + neutrofilos.c + linfocitos.c + nlr.c +
    platelets.c + saturacion_de_oxigeno.c + fio2.c + pafi.c + pao2.c +
    corticoides + anticoagulantes + antipaludicos + pronacion,
  data = data_uv_dm,
  family = binomial(link = "logit")
)

# Visual check of model assumptions
performance::check_model(m1_dm)

# Model performance
broom::glance(m1_dm)

# Indices of model performance
performance::model_performance(m1_dm)

# Check for multicollinearity
performance::check_collinearity(m1_dm)
```

```{r include=FALSE}
# Model summaries
epiDisplay::logistic.display(m1_dm, simplified = FALSE, crude = FALSE)
```

#### Complete multivariable for patients without T2DM

```{r}
# Non-T2DM model
full_multivariable_nondm <- glm(
  outcome ~
    edad.c + sex + hta + obesity + fever + tos + taquipnea + disnea +
    frecuencia_cardiaca.c + frecuencia_respiratoria.c + p_a_sistolica.c +
    p_a_diastolica.c + wbc.c + neutrofilos.c + linfocitos.c + nlr.c +
    platelets.c + saturacion_de_oxigeno.c + fio2.c + pafi.c + pao2.c +
    corticoides + anticoagulantes + antipaludicos + pronacion,
  data = data_uv_nondm,
  family = binomial(link = "logit")
) |>
  tbl_regression(
    exponentiate = TRUE,
    conf.int = TRUE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      sex ~ "Sex",
      hta ~ "Hypertension",
      obesity ~ "Obesity",
      fever ~ "Fever",
      tos ~ "Dry cought",
      taquipnea ~ "Tachypnea",
      disnea ~ "Dyspnea",
      frecuencia_cardiaca.c ~ "Heart rate (BPM)",
      frecuencia_respiratoria.c ~ "Respiratory rate (BPM)",
      p_a_sistolica.c ~ "SBP (mmHg)",
      p_a_diastolica.c ~ "DBP (mmHg)",
      wbc.c ~ "White-cells (×10^−9^/L)",
      neutrofilos.c ~ "Neutrophils (×10^−9^/L)",
      linfocitos.c ~ "Lymphocytes (×10^−9^/L)",
      nlr.c ~ "NLR",
      platelets.c ~ "Platelets (×10^−9^/L)",
      saturacion_de_oxigeno.c ~ "SaO~2~",
      fio2.c ~ "FiO~2~ (%)",
      pafi.c ~ "PaO~2~:FiO~2~ ratio",
      pao2.c ~ "PaO~2~",
      corticoides ~ "Corticosteroids",
      anticoagulantes ~ "Anticoagulants",
      antipaludicos ~ "Antimalarials",
      pronacion ~ "Pronation"
    )
  ) |>
  bold_p(t = 0.05) |>
  add_vif() |>
  modify_header(estimate = "**Multivariable OR**",
                p.value = "**p value**",
                aGVIF ~ "**aGVIF**")
```

```{r fig.height=10, fig.width=10, fig.align='center'}
m1_non = glm(
  outcome ~
    edad.c + sex + hta + obesity + fever + tos + taquipnea + disnea +
    frecuencia_cardiaca.c + frecuencia_respiratoria.c + p_a_sistolica.c +
    p_a_diastolica.c + wbc.c + neutrofilos.c + linfocitos.c + nlr.c +
    platelets.c + saturacion_de_oxigeno.c + fio2.c + pafi.c + pao2.c +
    corticoides + anticoagulantes + antipaludicos + pronacion,
  data = data_uv_nondm,
  family = binomial(link = "logit")
)

# Visual check of model assumptions
performance::check_model(m1_non)

# Model performance
broom::glance(m1_non)

# Indices of model performance
performance::model_performance(m1_non)

# Check for multicollinearity
performance::check_collinearity(m1_non)
```

```{r include=FALSE}
# Model summaries
epiDisplay::logistic.display(m1_non, simplified = FALSE, crude = FALSE)
```

```{r}
# Stack tables
table_mv = tbl_merge(
  tbls = list(
    table_s1.1_uv,
    full_multivariable_dm,
    table_s1.2_uv,
    full_multivariable_nondm
  )
) |>
  modify_spanning_header(
    c(estimate_1:aGVIF_2) ~ "**Patients with T2DM**",
    c(estimate_3:aGVIF_4) ~ "**Patients without T2DM**"
  ) |> 
  modify_footnote(everything() ~ NA, abbreviation = TRUE) |>
  modify_table_body( ~ .x |> arrange(row_type == "glance_statistic")) |>
  modify_footnote(
    c(estimate_1, estimate_2, estimate_3, estimate_4) ~ "OR = Odds Ratio",
    c(ci_1, ci_2, ci_3, ci_4) ~ "95% CI = 95% Confidence Interval",
    c(GVIF_2, GVIF_4) ~ "GVIF = Generalized Variance Inflation Factor",
    c(aGVIF_2, aGVIF_4) ~ "aGVIF = Adjusted GVIF or GVIF^[1/(2*df)]"
  )

# Convert gtsummary object to a flextable object and format character columns
flex_table_mv <- table_mv |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
flex_table_mv
```
#### Final model for patients with T2DM

```{r}
table_5.1_uv <- data_uv_dm |>
  tbl_uvregression(
    include = c(
      edad.c,
      hta,
      frecuencia_cardiaca,
      nlr,
      creatinine,
      saturacion_de_oxigeno,
      pao2,
      antipaludicos
    ),
    method = coxph,
    y = Surv(len_hosp_stay, outcome_ph),
    exponentiate = TRUE,
    conf.int = TRUE,
    hide_n = TRUE,
    add_estimate_to_reference_rows = FALSE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      hta ~ "Hypertension",
      frecuencia_cardiaca ~ "Heart rate (BPM)",
      nlr ~ "NLR",
      creatinine ~ "Creatinine (mg/dL)",
      saturacion_de_oxigeno ~ "SaO~2~",
      pao2 ~ "PaO~2~",
      antipaludicos ~ "Antimalarials"
    )
  ) |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Univariate HR**", p.value = "**p value**")
```

```{r}
table_5.1_mv <-
  coxph(
    Surv(len_hosp_stay, outcome_ph) ~ edad.c + hta + frecuencia_cardiaca + 
      nlr + saturacion_de_oxigeno + creatinine + saturacion_de_oxigeno +
      pao2 + antipaludicos,
    data = data_uv_dm
  ) |>
  tbl_regression(
    conf.int = TRUE,
    exponentiate = TRUE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      hta ~ "Hypertension",
      frecuencia_cardiaca ~ "Heart rate (BPM)",
      nlr ~ "NLR",
      creatinine ~ "Creatinine (mg/dL)",
      saturacion_de_oxigeno ~ "SaO~2~",
      pao2 ~ "PaO~2~",
      antipaludicos ~ "Antimalarials"
    )
  ) |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Multivariable HR**", p.value = "**p value** ")
```

#### Final model for patients without T2DM

```{r}
table_5.2_uv <- data_uv_nondm |>
  tbl_uvregression(
    include = c(
      edad.c,
      hta,
      frecuencia_cardiaca,
      nlr,
      saturacion_de_oxigeno,
      creatinine,
      saturacion_de_oxigeno,
      pao2,
      antipaludicos
    ),
    y = outcome,
    method = glm,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    conf.int = TRUE,
    hide_n = TRUE,
    add_estimate_to_reference_rows = FALSE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      hta ~ "Hypertension",
      frecuencia_cardiaca ~ "Heart rate (BPM)",
      nlr ~ "NLR",
      creatinine ~ "Creatinine (mg/dL)",
      saturacion_de_oxigeno ~ "SaO~2~",
      pao2 ~ "PaO~2~",
      antipaludicos ~ "Antimalarials"
    )
  ) |>
  bold_labels() |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Univariate HR**", p.value = "**p value**")
```

```{r}
table_5.2_mv <-
  coxph(
    Surv(len_hosp_stay, outcome_ph) ~ edad.c + hta + frecuencia_cardiaca + 
      nlr + creatinine + saturacion_de_oxigeno + pao2 + antipaludicos,
    data = data_uv_nondm
  ) |>
  tbl_regression(
    conf.int = TRUE,
    exponentiate = TRUE,
    pvalue_fun = ~ style_pvalue(.x, digits = 3),
    estimate_fun = ~ style_number(.x, digits = 2),
    label = list(
      edad.c ~ "Age (years)",
      hta ~ "Hypertension",
      frecuencia_cardiaca ~ "Heart rate (BPM)",
      nlr ~ "NLR",
      creatinine ~ "Creatinine (mg/dL)",
      saturacion_de_oxigeno ~ "SaO~2~",
      pao2 ~ "PaO~2~",
      antipaludicos ~ "Antimalarials"
    )
  ) |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Multivariable HR**", p.value = "**p value** ")
```

```{r}
# Number of observations
count.dm <- nrow(data_uv_dm)
count.non <- nrow(data_uv_nondm)

# Merge tables
table_5 <-
  tbl_merge(tbls = list(table_5.1_uv, table_5.1_mv,
                        table_5.2_uv, table_5.2_mv)) |>
  modify_spanning_header(list(
    c(estimate_1:p.value_2) ~ "**Patients with T2DM** (N = {paste(count.dm)})",
    c(estimate_3:p.value_4) ~ "**Patients without T2DM** (N = {paste(count.non)})"
  )) |>
  modify_footnote(everything() ~ NA, abbreviation = TRUE) |>
  modify_footnote(
    c(estimate_1, estimate_2, estimate_3, estimate_4) ~ "HR = Hazard Ratio",
    c(ci_1, ci_2, ci_3, ci_4) ~ "95% CI = 95% Confidence Interval"
  ) |>
  modify_caption("Table 5. Multivariable Cox Regression by T2DM")

# Convert gtsummary object to a flextable object and format character columns
flex_table_5 <- table_5 |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
flex_table_5
```

## Multiple Imputation with MICE
To adapt
https://stackoverflow.com/questions/53014141/mice-splitting-imputed-data-for-further-analysis
For descriptive

```{r}
# Selecction of variables that may help to impute variables with missing data
input_data =
  data_matched |>
    dplyr::select(
      # Demographic characteristics and clinical history
      edad,
      edad.c,
      sex,
      t2dm,
      hta,
      # Signs and symptoms
      fever,
      tos,
      dolor_de_garganta,
      malestar_general,
      taquipnea,
      disnea,
      # Vital signs
      frecuencia_respiratoria,
      frecuencia_cardiaca,
      p_a_sistolica,
      p_a_diastolica,
      # Laboratory findings
      wbc,
      neutrofilos,
      linfocitos,
      nlr,
      platelets,
      glucose,
      urea,
      creatinine,
      # Blood gas findings
      saturacion_de_oxigeno,
      fio2,
      pafi,
      pao2,
      # Treatments
      antipaludicos,
      # Follow-up
      len_hosp_stay,
      # outcomes
      outcome_ph
    )
```

### Missing data pattern

```{r}
colSums(is.na(input_data))

data_matched |>
  dplyr::select(
    # Vital signs
    frecuencia_respiratoria,
    frecuencia_cardiaca,
    p_a_sistolica,
    p_a_diastolica,
    # Laboratory findings
    wbc,
    neutrofilos,
    linfocitos,
    nlr,
    platelets,
    glucose,
    urea,
    creatinine,
    # Blood gas findings
    saturacion_de_oxigeno,
    fio2,
    pafi,
    pao2,
    # Follow-up
    len_hosp_stay
    ) |>
  ggmice::plot_pattern(
    square = TRUE,
    rotate = TRUE,
    #npat = 3,
    #cluster = "reg"
  )
```
### Missing vs no missing

```{r}
table_missing <- data_matched |> 
  dplyr::select(
    # Vital signs
    frecuencia_respiratoria,
    frecuencia_cardiaca,
    p_a_sistolica,
    p_a_diastolica,
    # Laboratory findings
    wbc,
    neutrofilos,
    linfocitos,
    nlr,
    platelets,
    glucose,
    urea,
    creatinine,
    # Blood gas findings
    saturacion_de_oxigeno,
    fio2,
    pafi,
    pao2,
    # Follow-up
    len_hosp_stay
    ) |>
  mutate(missing = factor(
    is.na(urea),
    levels = c(FALSE, TRUE),
    labels = c("Not Missing", "Missing")
  )) |> 
  tbl_summary(
    by = missing,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n}    ({p}%)")
    ) |> 
  modify_header(label = "**Variable**",
                all_stat_cols() ~ "**{level}**<br>N = {n} ({style_percent(p, digits=1)}%)") |> 
  modify_caption("Participant characteristics, by missingness") |> 
  bold_labels()
```

### Multiple data imputation, m = 5

```{r}
## Imputation with mice
data_imputated =
  mice(
    input_data,
    m = 5,
    method = c(
      # For demographics
      "",
      "",
      "",
      "",
      "",
      # Signs and Symptoms
      "",
      "",
      "",
      "",
      "",
      "",
      # Vital signs
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      # Laboratory findings
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      # Blood gas findings
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      # Treatments
      "",
      # Follow up and outcomes
      "pmm",
      ""),
    maxit = 20,
    seed = 500
    )

data_imputated_l <- complete(data_imputated, "long", include = TRUE)
```


```{r}
ggmice(data_imputated, aes(x = .imp, y = glucose)) +
  geom_jitter(height = 0, width = 0.25) +
  geom_boxplot(width = 0.5, size = 1, alpha = 0.55, outlier.shape = NA) +
  labs(x = "Imputation number")

purrr::map(c("glucose", "creatinine", "urea"), ~ {
  ggmice(data_imputated, aes(x = .imp, y = .data[[.x]])) +
    geom_boxplot() +
    labs(x = "Imputation number")
}) |> 
  patchwork::wrap_plots()
```

### Multiple data imputation, m = 20

```{r}
## Imputation with mice
data_imputated =
  mice(
    input_data,
    m = 20,
    method = c(
      # For demographics
      "",
      "",
      "",
      "",
      "",
      # Signs and Symptoms
      "",
      "",
      "",
      "",
      "",
      "",
      # Vital signs
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      # Laboratory findings
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      # Blood gas findings
      "pmm",
      "pmm",
      "pmm",
      "pmm",
      # Treatments
      "",
      # Follow up and outcomes
      "pmm",
      ""),
    maxit = 20,
    seed = 500
    )
```

### Adapt for design: separate for patients with diabetes and without diabetes

```{r}
d.long <- mice::complete(data_imputated, "long", include = T)

d.long.dm <- d.long[which(d.long$t2dm == 'yes'),]
d.long.nondm <- d.long[which(d.long$t2dm == 'no'),]

data_imputated_dm <- as.mids(d.long.dm)
data_imputated_nondm <- as.mids(d.long.nondm)

data_imputated_dm$imp$creatinine
data_imputated_nondm$imp$creatinine
```
###

```{r fig.height=10, fig.width=10, fig.align='center'}
model_imputation_dm <-
  with(data_imputated_dm, coxph(
    Surv(len_hosp_stay, outcome_ph) ~ edad.c + hta + frecuencia_cardiaca + 
      nlr + creatinine + saturacion_de_oxigeno + pao2 + antipaludicos)
    )

model_imputation_nondm <-
  with(data_imputated_nondm, coxph(
    Surv(len_hosp_stay, outcome_ph) ~ edad.c + hta + frecuencia_cardiaca + 
      nlr + creatinine + saturacion_de_oxigeno + pao2 + antipaludicos)
    )

pool.imputation_dm <- pool(model_imputation_dm)

summary_results = summary(pool.imputation_dm, conf.int = TRUE)
```

```{r}
##
tbl_dm_imput <-
  data_imputated_dm |> 
  with(coxph(Surv(len_hosp_stay, outcome_ph) ~ edad.c + 
                                     hta + frecuencia_cardiaca + nlr + 
                                     creatinine + saturacion_de_oxigeno + 
                                     pao2 + antipaludicos)) |> 
  tbl_regression(exponentiate = TRUE,
                 label = list(
                   edad.c ~ "Age (years)",
                   hta ~ "Hypertension",
                   frecuencia_cardiaca ~ "Heart rate (BPM)",
                   nlr ~ "NLR",
                   creatinine ~ "Creatinine (mg/dL)",
                   saturacion_de_oxigeno ~ "SaO~2~",
                   pao2 ~ "PaO~2~",
                   antipaludicos ~ "Antimalarials")) |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Multivariable HR**", p.value = "**p value** ")
  
tbl_nondm_imput <-
  data_imputated_nondm |> 
  with(coxph(Surv(len_hosp_stay, outcome_ph) ~ edad.c + 
                                     hta + frecuencia_cardiaca + nlr + 
                                     creatinine + saturacion_de_oxigeno + 
                                     pao2 + antipaludicos)) |> 
  tbl_regression(exponentiate = TRUE,
                 label = list(
                   edad.c ~ "Age (years)",
                   hta ~ "Hypertension",
                   frecuencia_cardiaca ~ "Heart rate (BPM)",
                   nlr ~ "NLR",
                   creatinine ~ "Creatinine (mg/dL)",
                   saturacion_de_oxigeno ~ "SaO~2~",
                   pao2 ~ "PaO~2~",
                   antipaludicos ~ "Antimalarials")) |>
  bold_p(t = 0.05) |>
  modify_header(estimate = "**Multivariable HR**", p.value = "**p value** ")
```
### Merge tables

```{r}
# Merge tables
table_5_imput = tbl_merge(
  list(tbl_dm_imput, tbl_nondm_imput),
  tab_spanner = c("**Patients with T2DM**", "**Patients without T2DM**"
  )) |>
  modify_caption("Table S1. Multivariable Cox regression by T2DM with Multiple Imputatated Data")

# Convert gtsummary object to a flextable object and format character columns
flex_table_s1 <- table_uv |>
  gtsummary::as_flex_table() |>
  my_flextable_theme() |>
  flextable::fontsize(part = "all", size = 10) |>
  ftExtra::colformat_md()

# View
flex_table_s1
```



